Closing the gap between statistical and scientific workflows for improved forecasts in ecology

### 2025-09-19 revision/r2/response_r2.tex

We thank the reviewer for their positive comments regarding our manuscript. 
We have worked to correct these inconsistencies.

{\bf Reviewer 2}

\begin{mybox}
\emph{All of my major comments were sufficiently addressed in this revision of
the manuscript. I appreciate that the supplement is now referenced in
the main document (I missed it on my last round of review). I have a few
editorial comments for the main text and a few minor comments about the
supplement, but I feel this manuscript is almost ready for publication.}
\end{mybox}

We are glad that the reviewer found the manuscript has improved.
Thank you for this last feedback.

@@

\begin{mybox}
\emph{Line 215: comma after parameters}
\end{mybox}

Added.

\begin{mybox}
\emph{A couple typos in Box e.g., "calibrating submodes separately allows to
avoid" missing "us"?}
\end{mybox}

Changed to: ``Calibrating submodels separately makes possible to avoid the fact that [...]''

\begin{mybox}
\emph{\textbf{Supplement Minor comments:}\\
Although the supplement works through an example far simpler than the
types of research questions discussed in the main text, it serves as a
reasonable (and accessible) example of how the workflow could be
implemented. There is a brief discussion towards the end of the
supplement about the ``scalability" of the proposed workflow, but it
leaves something to be desired. Have the authors applied this workflow
(or pieces of it) to examples that are more complex that could be cited,
or is that future work?}
\end{mybox}

@@

...

\begin{mybox}
\emph{First green box: typo in RQ..``What is global trend..." should read
``what is the global trend..."}
\end{mybox}

@@

\begin{mybox}
\emph{Stan model not included so could not be assessed (model1\_nc2.stan), I
recommend sharing for transparency.}
\end{mybox}

Sorry we forgot to share the Stan code. We added the code in the workflow example.

\begin{mybox}
\emph{Figure 15, unclear what the posterior predictive distribution is of,
this detail could be added for transparency. Could also add many
posterior predictive datasets to the same plot and then overlay the
observed data to facilitate easier comparisons between y-tilde and y-obs.}
\end{mybox}

% Hmm I don't understand?

{\bf Reviewer 3}

\begin{mybox}
\emph{First, I would like to congratulate the authors for a very clear and well written manuscript. The main thesis of the authors is to outline modelling workflows and illustrate in broad brush strikes how to bridge the current gap in ecological forecasting with simulations of synthetic data among the several steps of model building (the pre-data/post-model third corner in their Fig. 1) to improve confidence in modelling, inference and prediction. This is an important contribution, one that echoes many early calls (e.g. Thomas 1997): our community in ecology, despite its beefing up with respect to quantitative skills, remains fragmented and disjoint. I thus wholeheartedly endorse the main thesis of the authors and agree with it about research. There remains one concern that I would like to discuss with the authors, and possibly hear back from them although it is somewhat minor with respect to research but more problematic with respect to policy. It should not deflect from publication of the article in Philosophical Transactions of the Royal Society A.}
\end{mybox}

We are glad that the reviewer found our manuscript well-written, and that he endorses our main messages. We are particularly grateful for his detailed comments, with much broader perspectives than just our manuscript. We have tried to address most of the comments, but we believe that they deserve longer discussions and could constitute in themselves another paper.

\begin{mybox}
\emph{The authors have addressed comments from the first round of reviewing. Figure 1 is central to their thesis: it is clear and well explained. However, I feel the lower left corner, corresponding to the pre-model/post-data area should be some darker grey to highlight that this corner is not overlooked but simply does not make a lot of sense in the current proposal (and in the idealized view of the current scientific process). It may sometimes correspond to mindless monitoring if one is uncharitable (Yoccoz et al. 2001, Legga \& Nagy 2006). I believe it should not be greyed out the same colour as the other corners.}  
\end{mybox}

\begin{mybox}
\emph{Moving to the upper right corner (post-model/pre-data), the authors could mention
Simulation-Based Calibration (Modrák et al. 2025) to point readers to this useful tool for code
checking, which is also central to workflows. I suspect this omission might be wilful not to
overload readers. It might still be useful to point SBC on ecologists’ map as new models keep
being proposed but are insufficiently put to the test beyond one case study in many instances
(see also Nichols et al. 2019).}  
\end{mybox}

Thank you for this suggestion. We now mention simulation-based calibration (\href{file:forecastflows_r2\#lintarget:SBC}{line \ref*{lin:SBC}}).

\begin{mybox}
\emph{With respect to terminology, the authors are mentioning “research spheres” in the abstract,
which is a catchy term, and one I agree may be more appropriate than that of a research
programme for what the authors are discussing. What is a research sphere is not articulated
further by the authors and there must be a “policy sphere” at which the paper is alluding to. In
fact, where I find the authors’ thesis is less convincing is with respect to policy outreach. In
policy, vague terminology such a “research sphere” is useful to get discussions started as
loosely defined terms facilitate interactions. This can also lead to talking past each other but,
usually, terminology gets to be clarified at some point when misunderstandings are too
numerous to be ignored (and one get a common understanding document out of a round of
meetings). My point is that the expectations of the “policy sphere” are usually not completely
aligned with the “research sphere”. And this realization bears on workflows.}  
\end{mybox}

\begin{mybox}
\emph{In policy, a trend in abundance is needed as evidence for decision and further actions. Whether
that trend comes from a phenomenological or mechanistic approach is often of secondary
interest to policy-makers: what matters is that the trend passes some checks (including
statistical significance). Much has been written on the topic, especially with respect to
statistical power. So “outlin[ing] the steps of a universal workflow” (line 48-49) is both novel
and worthy. Yet I am not convinced that it will solve some of issues that are pregnant in the
policy sphere. For example, timeliness is important in policy. As the authors are right to point,
workflows require time to be developed. Sometimes a lot more time is also needed just to agree
on the workflow, and then there is time and man-power needed for documenting, maintaining
and updating the workflow. In short, one need a governance structure for the workflow. If the
goal is to evidence a negative trend because this is the minimum requirement for taking
corrective actions, then the authors’ proposal, while clearly an improvement for the research
sphere, will not impress the policy sphere as it is also likely to deliver too late; nor will it
necessarily lead to better decisions. Awareness, design and publication of workflows will also
have to happen in relevant policy circles, not just in scientific journals. That we should start
with scientific journals in the right first steps, but reaching out to the policy sphere will not
happen on its own either.}  
\end{mybox}

\begin{mybox}
\emph{I disagree with the framing on lines 7-9: the explanation provided from the two paths is not
very convincing. The authors seem to have implicitly in mind some global policy, and hence
justify a need for a global analysis. However, much of conservation is local, national or
regional. Some surveys are designed and carried out at these scales, but when pooled in a
global analysis, their analysis can become problematic because of sampling imbalances.
However, the aim is now to make a global claim but that does not mean the local or regional
trend was inaccurate in the first place. Yet the global analysis may be taken as evidence of
biases (more on this term below) in policy fora. I overall think there is an unspoken assumption
throughout in this opinion piece which is that all policies are somehow global. I see here a
disconnect between high-profile papers on global trends which get attention precisely because
they are global (and likely to get many citations and thus are a staple of high impact journals
too), and a lot of policy which is much more local in scale.}  
\end{mybox}

We do agree that our framing is not directly adapted to the more local-scale decisions. We now mention that the policy needs we refered to are rather ``global'' (\href{file:forecastflows_r2\#lintarget:global}{line \ref*{lin:global}}).

\begin{mybox}
\emph{To be clearer, I will provide an example (see Authier 2025). Using citizen science data on 724
species, Callaghan et al. (2021) estimated global population sizes for 9,700 extant bird species
with modelling. These model-based estimates are not gauged against independent estimates
for species well-monitored at the global scale. Owing to their large-size and conspicuousness
several seabirds species can be censused accurately. Although imperfect, these estimates
represent the best available knowledge for policy (ACAP 2021). Model-based abundances
were > 20 times higher/smaller than IUCN ones for 45/5 (out of 96 seabird species) and up to
100 times higher/smaller for 18/1 species (Figure 1). This is particularly concerning for
endemic species, in which the local and global geographical scales are completely confounded,
and which can comprise a single small population. A startling example is the Mascarene petrel
(Pseudobulweria aterrima), a marine megafauna species endemic to Reunion Island in the Indian
Ocean. Data on Mascarene petrels were used in the model training set and the estimated
abundance by Callaghan et al. (2021) was several thousands of individuals. The estimated
population size from local scientists is less than a few hundred individuals (Lopez et al. 2021).
The global analysis is not helpful in my opinion, and may even hinder local policy which have
taken a long time to gain attention and momentum.} 
\end{mybox}

This is a very interesting example. As highlighted in your next comment, maybe Callaghan et al. would have identified some flaws in their modeling approach if they had spent more time in the pre-data/post-model quadrat? If they predicted abundance estimations for 9000 new species outside their data, it seems like their inferential goal was above all prediction---and thus they should have focused on testing this goal in their workflow and should have developed some data simulation steps coherent with this inferential goal. It would be a very interesting example applied to conservation to illustrate how step-by-step workflows could help identify these local-to-global discrepancies, and maybe \emph{in fine} reconcile local observations and actions with global trends and policies?

\begin{mybox}
\emph{Callaghan et al. (2021) are concerned about forking paths and aware of reproducibility issues.
They described a validation workflow which does not take stock of the pre-data/post-model
corner but claimed nevertheless accuracy and success in redressing some biases (e.g.
geographic and taxonomic). More generally, this example illustrates the following science-
policy workflow: policy-makers read about a new global analysis published in a scientific
journal. The analysis has a model workflow. Policy-makers look for the modelled-based
abundance of their pet species and compare those to their knowledge and data. Local, national
or regional scientists involved in policy are then contacted to explain any discrepancies. If
there are indeed discrepancies, the scientists have now to explain their origins, looking into the
model workflow and how it espoused good practices or not. With a growing familiarity of
what are workflows, this should become more and more seamless in the long run. In the short
run however, this science-policy workflow (i.e. the linear model of science, see below) is not
scalable with respect to scientist’s time in my opinion and it will not necessarily help policy if
scientists (especially non-tenured ones) think that this is time spent not doing field work,
writing grants and supervising students to further their research.}  
\end{mybox}

First, we would like to answer about the time it may take for this type of workflow to become more common in research spheres. A large portion of scientific results today are investigated and written by non-tenured scientists. We have examples around us of graduate students who are likely more inclined than their supervisors to spend time developing carefully checked models and following workflow steps that give them more confidence in their results. While their supervisors may not push them to spend more time doing so, they will not prevent them from doing it either. Therefore, we believe that this paper---among many others---could actually help spread the use of robust modeling workflows (faster than it seems). Although the road between being a graduate student and obtaining a tenure-track position is long and uncertain, those graduate students are soon to become reviewers and submit grants (if they decide to pursue in academia), and interact with decision-makers, at least at a local scale (and even for those who will quit academia).

\begin{mybox}
\emph{The above scenario (which can be applied to global analyses with or without workflows) is not
to deflect from the authors’ main point being made that “that an improved workflow that
required data simulation and retrodictive checks would lead to larger model advances and a
greater recognition of uncertainty—thus highlighting likely consistency in estimates across
models—that could better aid policy”. The “could better aid policy” part is simply too vague
and could betray, in my opinion, that the proposed “universal framework” is implicitly
embracing the linear model of science (Fernándes 2016) of ‘truth speaks to power’. In other
words, bad policy decisions are assumed to mainly result from a deficit of knowledge so that
all that is needed would be better knowledge to naturally lead to good policy decision. I find
this less and less convincing given that we’ve never had so much data or knowledge than now.
We know what are the main drivers of biodiversity loss (e.g. Maxwell et al. 2016). We are
literally awash with models (just think about the number of Species Distribution Models being
published nowadays). Yet biodiversity, on average, does not fare any better.}  
\end{mybox}

\begin{mybox}
\emph{While I agree with the authors that the workflow will help research spheres in disciplining
modelling, I disagree with the policy implications. As highlighted in the box by the authors,
“[m]odel development is the central step of the process-based workflow, typically requiring
several years, yet it often remains opaque for anyone who has not worked the model itself.”
This is not a small issue.}  
\end{mybox}

We wanted to highlight that process-based model development can sometimes be very long (with or without this workflow), and thus that it would be better to follow coherent steps in model development and checking---given the amount of time process-based model development takes anyway. But to be clear, for most applications, applying the workflow do not require years.

\begin{mybox}
\emph{For example, in the International Whaling Commission, established in
1946, many scientists are now retiring without new ones stepping in their shoes at the Scientific
Committee (SC). The IWC SC pioneered some work on simulations (e.g. De la Mare 1986) and
devised complex population models that were heavily tested to assess whether they would be
fit-for-purposes. It is ironic that the simulation-savvy and model-based procedures
encapsulated in the Revised Management Procedure that was developed in the late 1980s and
early 1990s was never used in practice (for policy reasons, not scientific ones, to make a long
story short).}  
\end{mybox}

\begin{mybox}
\emph{The IWC example underlines the long-term commitment that is needed and the bigger
question is how to square such a commitment in a competitive academic landscape where
projects, when they are funded, are meant for a couple of years and the research proletariat is
expected to publish regularly in prestigious journals and be highly mobile with respect to
institutions. I contend that the incentives for the research and policy spheres are simply not
aligned enough at the moment. How does one ensure the sort of long-term commitment
needed in policy where stability of procedures is important (as any management procedures or
indicators are going to have to jump many hoops to get agreed by all parties) and a steady
stream of publications in an academic ecosystem where novelty and innovations are a
premium?}  
\end{mybox}

\begin{mybox}
\emph{The authors are alluding to the current publishing culture as a barrier but do not elaborate
much. That is fair to omit in a short paper, and again I agree with their main point with respect to the research spheres.}
\end{mybox}

\begin{mybox}
\emph{I do think, however, that evidencing a trend serves a different goal in
policy, and that having a workflow to bridge trend estimation and forecasting is not as urgent
as the authors imply if the ultimate aim is to enable timely decisions to address biodiversity
loss. In fact, the insistence on workflow may backfire simply because it can easily lead to either
distraction or displacement in the sense of Rainer (2012). Distraction may happen because it
could be easy to ask for more testing of some model features (a new twist to some mechanistic
parts in the model) and ask that no decision be taken before that an updated workflow is
completed. This could easily align with the common policy requirement of using the ‘best
available evidence’, especially if some cutting-edge new results are brought forward at the
policy table and one is being tasked with updating the current model. Displacement is more
subtle but can also happen: managing the workflow becomes the policy and simply
maintaining it absorb most of the available resources. As Rayner (2012) puts it “displacement
occurs when an organization engages with an issue, but substitutes management of a
representation of a problem (such as a computer model) for management of the represented
object or activity.” I think these two issues are real.} 
\end{mybox}

This is a very interesting question. However, we believe that the previous example of Callaghan et al. illustrates how bridging the differential goals of understanding and forecasting (or extrapolating to poorly observed species) can actually have important implications for policy. Similarly, it does not seem entirely accurate to think that this type of workflow should have been applied earlier (Revised Management Procedure of IWC) while also considering that it might be now 'too late' to highlight its importance (because there is no more time to lose to adress biodiversity loss?). We think this issue is linked to other challenges we encouter as scientists today: is there a risk in better quantifying and communicating the uncertainties of our results, if these uncertainties lead decision-makers to avoid making decisions? Or, by being the most transparent about our uncertain results (and spending more time applying consistent workflows), do we rather serve not only research but also the policy sphere?

\begin{mybox}
\emph{I would love to actually read the authors’ thoughts here and ideally be proven wrong that an
improved workflow would be enough for enabling policy decisions whose beneficial impacts
would be commensurate with the current magnitude of biodiversity loss7. In particular, I
would be interested in a better articulation of how they see global studies feed-backing on
policy which is usually more local, national or regional.}  
\end{mybox}

\begin{mybox}
\emph{My remaining points are more concise.
- Terminology. If the authors are also aiming at the policy sphere, the word bias (e.g. lines
13, 134) should be wielded with extreme caution as it is easily weaponized. In policy
circles, the term bias is, in my experience, the kiss of death as it is often interpreted as
lopsided evidence. The authors are discussing preferential sampling with respect to space,
time or taxonomy and its consequences in modelling results. Preferential sampling is more
descriptive than bias, which also has a prescriptive meaning, esp. in policy8. Also, a well-
designed survey at some spatial scale may become an instance of preferential geographic
sampling when collated with other surveys and data. This sort of data collation is often the
case with global trend analyses that the authors seem to have in mind. Also, taxonomic
biases are bound to happen in global analyses. How do we redress those? That is, what can
we say about data-poor species? How a richly-parametrized workflow calibrated on data
collected on a few very well-studied species is going to help the majority of data-poor
species? Taxonomic gaps have been identified without workflows and how to address
those gaps point to monitoring and the need of appropriate governance to ensure long-
term monitoring and inclusion of data-poor species. Research alone may not provide
enough incentive or impetus for inclusion of data-poor species, especially if what is
needed for policy about them is descriptive and routine science.}  
\end{mybox}

\begin{mybox}
\emph{- ML enthusiasm (lines 298-303): the point is granted for prediction but again evidencing a
trend for policy is often required for further action. I fail to see how ML, if focused on
prediction, is going to help. A trend is a useful fiction that is easily understood in policy
fora where interpretability still has a huge premium. And a trend is very much
interpretable. I think the authors may be going overboard in their claims if, again, their
proposal is to have policy-relevance and not only research-relevance (a worthy goal of its
own).}  
\end{mybox}

\begin{mybox}
\emph{- Robustness is great. Policy loves the word but too rarely elaborates on it. Thus, on line 62,
these methods should be robust against what more precisely?}  
\end{mybox}

\begin{mybox}
\emph{- typo on line 138: “These feedback loops are applied currently in only in Bayesian
frameworks (to our knowledge)...”}
\end{mybox}

### 2025-09-19 revision/r2/forecastflows_r2.tex
starting to answer the very interesting review of referee 3

Meeting these policy needs has led often to two separate paths: one focused on estimating trends from new global datasets, and another focused on forecasting from generally distinct datasets or mechanistic models based on fewer data. 
Newly available large-scale, long-term datasets have provided our first `global' estimates of biodiversity trends \citep[e.g.][]{loh2005living,Dornelas2018}, but these data---gathered opportunistically from multiple sources---are unbalanced and suffer from large geographic, temporal and taxonomic biases. Models to date have failed to fully address these challenges and, perhaps because of these limitations, are rarely if ever used for forecasting.
Instead, forecasting---under different plausible scenarios---has generally relied on entirely different datasets combined with either correlative or process-based models \citep{IPBES2019}, with process-based models often promoted as the most realistic approach \citep{Urban2016, Pilowsky2022} because they focus on mechanistic representations of ecosystem functioning. These approaches have failed to yield agreement on current species trends, leading to ongoing debates about the magnitude and even direction \citep{Dornelas2014, Leung2020, Buschke2021, Johnson2024}, and producing forecasts that diverge due to high model uncertainty at the ecological level \citep{Cheaib2012, Thuiller2019}.

@@

With a model in place, the next step focuses on testing and understanding it via data simulation (step 3, Figure \ref{fig:workflow}). `Fake' or `test' data are generated directly from the model by fixing parameters \llabel{firstsim}to some reasonable values, which is straightforward if the parameters are interpretable, and from `fake' (simulated) predictor data. 
We then fit our model to this simulated dataset and evaluate its ability to recover the prescribed values (\llabel{quad6}post-model/pre-data, Figure \ref{fig:modeldata}d). At this stage, the focus is on understanding the model, so we may need to spend time thinking about whether each parameter is realistic or not. This is also a step for making sure the model is working as expected (see example workflow). \llabel{repeatsim} Ideally, this data simulation step should be repeated several times with different parameter values, sampled within a reasonable range. \llabel{priorPC} In a Bayesian framework, simulated parameters can be sampled from the prior distribution (allowing us to check the implications of our priors \citealp{Gelman2020}), while in other frameworks they can be sampled from a range of biologically reasonable values (which is very similar to the concept of a prior).

\begin{figure}[h]
	\centering
	\vspace*{-0.5cm}
    \hspace*{-1.5cm}
	\includegraphics{../../figures/figure_worflow_wsteps}
	\caption{The workflow we propose here, which builds from recent advances in workflows \citep[][]{betanworkflow,Gelman2020,Schad2020,grinsztajn2021,vandeschoot2021,Wolkovich2024}, focuses on iterative feedbacks between the research question, model and data. Most research should start at Step 1, with the research question, followed by extensive model building and evaluation using simulated data (Steps 2 and 3) before proceeding to fitting the model on empirical data and examining it through retrodictive checks (Steps 4 and 5). Within this workflow, forecasting is a natural output of the process and not a separate process or one available for only certain modeling approaches.  The integration of observations (the data stream) occurs at Step 4---only after the model has been thoroughly evaluated---and can also highlight opportunities to collect new data and enrich the data stream. This figure is expanded from Figure \ref{fig:modeldata}d.}
	\vspace*{-0cm}
	\label{fig:workflow}
\end{figure}

@@

\noindent % Process-based models are built on explicit mathematical equations to describe (supposedly causal) relationships between environmental drivers and ecological responses. 
% They also often incorporate empirical relationships, particularly when knowledge is incomplete or when some processes are intentionally omitted. Processes are often represented at different nested spatiotemporal scales, depending on the underlying assumptions. 
Model development is the central step of the process-based workflow, typically requiring several years, yet it often remains opaque for anyone who has not worked the model itself. The step of designing the model---translating knowledge and hypotheses into mathematical equations and parameters---is often blurred with the step of model calibration (or tuning), where parameter values are inferred. Models are often treated as an accumulation of multiple submodels, each governing one or several ecological processes (see Figure B above). Rather than being fitted as a whole, submodels are calibrated separately against specific subsets of data, and some parameters are simply prescribed (i.e., fixed to a value found in the literature) or tuned to reproduce certain observations or theory. The way models are currently calibrated is likely not a coincidence, but rather a workaround to avoid confronting the full complexity of the model. Calibrating submodels separately allows to avoid the fact that, if the model were fitted as a whole, many parameters would compensate for one another---revealing structural degeneracies and making the model far more difficult to use.  %emw24Jul: nice stuff here.

>>>>

Meeting these \llabel{global}global policy needs has led often to two separate paths: one focused on estimating trends from new global datasets, and another focused on forecasting from generally distinct datasets or mechanistic models based on fewer data. 
Newly available large-scale, long-term datasets have provided our first `global' estimates of biodiversity trends \citep[e.g.][]{loh2005living,Dornelas2018}, but these data---gathered opportunistically from multiple sources---are unbalanced and suffer from large geographic, temporal and taxonomic biases. Models to date have failed to fully address these challenges and, perhaps because of these limitations, are rarely if ever used for forecasting.
Instead, forecasting---under different plausible scenarios---has generally relied on entirely different datasets combined with either correlative or process-based models \citep{IPBES2019}, with process-based models often promoted as the most realistic approach \citep{Urban2016, Pilowsky2022} because they focus on mechanistic representations of ecosystem functioning. These approaches have failed to yield agreement on current species trends, leading to ongoing debates about the magnitude and even direction \citep{Dornelas2014, Leung2020, Buschke2021, Johnson2024}, and producing forecasts that diverge due to high model uncertainty at the ecological level \citep{Cheaib2012, Thuiller2019}.

@@

With a model in place, the next step focuses on testing and understanding it via data simulation (step 3, Figure \ref{fig:workflow}). `Fake' or `test' data are generated directly from the model by fixing parameters \llabel{firstsim}to some reasonable values, which is straightforward if the parameters are interpretable, and from `fake' (simulated) predictor data. 
We then fit our model to this simulated dataset and evaluate its ability to recover the prescribed values (\llabel{quad6}post-model/pre-data, Figure \ref{fig:modeldata}d). At this stage, the focus is on understanding the model, so we may need to spend time thinking about whether each parameter is realistic or not. This is also a step for making sure the model is working as expected (see example workflow). \llabel{repeatsim} Ideally, this data simulation step should be repeated several times with different parameter values, sampled within a reasonable range. \llabel{priorPC} In a Bayesian framework, simulated parameters can be sampled from the prior distribution (allowing us to check the implications of our priors \citealp{Gelman2020}, \llabel{SBC}and to validate model inferences using simulation-based calibration \citealp{Talts2018_SBC}). In other frameworks, they can be sampled from a range of biologically reasonable values (which is very similar to the concept of a prior).

\begin{figure}[h]
	\centering
	\vspace*{-0.4cm}
    \hspace*{-1.5cm}
	\includegraphics{../../figures/figure_worflow_wsteps}
	\caption{The workflow we propose here, which builds from recent advances in workflows \citep[][]{betanworkflow,Gelman2020,Schad2020,grinsztajn2021,vandeschoot2021,Wolkovich2024}, focuses on iterative feedbacks between the research question, model and data. Most research should start at Step 1, with the research question, followed by extensive model building and evaluation using simulated data (Steps 2 and 3) before proceeding to fitting the model on empirical data and examining it through retrodictive checks (Steps 4 and 5). Within this workflow, forecasting is a natural output of the process and not a separate process or one available for only certain modeling approaches.  The integration of observations (the data stream) occurs at Step 4---only after the model has been thoroughly evaluated---and can also highlight opportunities to collect new data and enrich the data stream. This figure is expanded from Figure \ref{fig:modeldata}d.}
	\vspace*{-0cm}
	\label{fig:workflow}
\end{figure}

@@

\noindent % Process-based models are built on explicit mathematical equations to describe (supposedly causal) relationships between environmental drivers and ecological responses. 
% They also often incorporate empirical relationships, particularly when knowledge is incomplete or when some processes are intentionally omitted. Processes are often represented at different nested spatiotemporal scales, depending on the underlying assumptions. 
Model development is the central step of the process-based workflow, typically requiring several years, yet it often remains opaque for anyone who has not worked the model itself. The step of designing the model---translating knowledge and hypotheses into mathematical equations and parameters---is often blurred with the step of model calibration (or tuning), where parameter values are inferred. Models are often treated as an accumulation of multiple submodels, each governing one or several ecological processes (see Figure B above). Rather than being fitted as a whole, submodels are calibrated separately against specific subsets of data, and some parameters are simply prescribed (i.e., fixed to a value found in the literature) or tuned to reproduce certain observations or theory. The way models are currently calibrated is likely not a coincidence, but rather a workaround to avoid confronting the full complexity of the model. Calibrating submodels separately makes possible to avoid the fact that, if the model were fitted as a whole, many parameters would compensate for one another---revealing structural degeneracies and making the model far more difficult to use.  %emw24Jul: nice stuff here.
